News Across Languages - Cross-Lingual Document Similarity and Event Tracking

### 2015-09-05 rupnik2015.tex
typos

\subsection{Problem definition}\label{sec:tfidf}
We will first describe how documents are represented as vectors and how to compare douments in a mono-lingual setting. We then define a way to measure cross-lingual similarity which is natural for the models we consider.

@@

\noindent\textbf {Cross-lingual similarity.}
Processing a multilingual dataset results in several vector spaces with varying dimensionality, one for each language. The dimensionality of the vector space corresponding to the $i$-th language is denoted by $n_i$ and and the vector space model mapping is denoted by $\phi_i : \text{text} \rightarrow \RR^{n_i}$.
The similarity between documents in language $i$ and language $j$ is defined as a bilinear operator represented as a matrix $S_{i,j} \in \RR^{n_i \times n_j}$:
$$sim_{i,j}(d_1, d_2) = \frac{ \langle \phi_i (d_1), S_{i,j} \phi_j (d_2) \rangle }{\|\phi_i(d_1)\| \|\phi_j(d_2)\|},$$
where $d_1$ and $d_2$ are documents written in the $i$-th and $j$-th language respectively. If the maximal singular value of $S_{i,j}$ is bounded by $1$, then the similarity scores will lie on the interval $[-1, 1]$. We will provide an overview of the models in Section \ref{sec:models} and then introduce additional notation in \ref{sec:notation}. Starting with Section \ref{sec:kmeans} and ending with Section \ref{sec:hublang} we will describe some approaches to computing $S_{i,j}$ given training data.

\subsection{Cross-Lingual Models}\label{sec:models}
In this section, we will describe several approaches to the problem of computing multilingual simliarities introduced in \ref{sec:tfidf}. We present four approaches:
a simple approach based on $k$-means clustering in \ref{sec:kmeans}, a standard approach based on singular value decomposition in \ref{sec:LSI}, a related
approach called Canonical Correlation Analysis (CCA) in \ref{sec:CCA} and finally a new method, which is an extension of CCA to more than two languages in \ref{sec:hublang}.
%
CCA can be used to find correlated patterns for a pair of languages, whereas the extended method optimizes a
Sum of Squared Correlations (SSCOR) between several language pairs, which was introduced in \cite{Kettenring}. The SSCOR problem is difficult to solve in our setting (hundreds of thousands of features, hundreds of thousands of examples). To tackle this, we propose a method which consists of two ingredients.
 The first one is based on an observation that certain datasets (such as Wikipedia) are biased towards one language (English), which can be exploited
 to reformulate a difficult optimization problem as an eigenvector problem. The second ingredient is dimensionality reduction using LSI, which
 makes the eigenvector problem computationally and numerically tractable.

>>>>

\subsection{Problem definition}\label{sec:tfidf}
We will first describe how documents are represented as vectors and how to compare documents in a mono-lingual setting. We then define a way to measure cross-lingual similarity which is natural for the models we consider.

@@

\noindent\textbf {Cross-lingual similarity.}
Processing a multilingual dataset results in several vector spaces with varying dimensionality, one for each language. The dimensionality of the vector space corresponding to the $i$-th language is denoted by $n_i$ and the vector space model mapping is denoted by $\phi_i : \text{text} \rightarrow \RR^{n_i}$.
The similarity between documents in language $i$ and language $j$ is defined as a bilinear operator represented as a matrix $S_{i,j} \in \RR^{n_i \times n_j}$:
$$sim_{i,j}(d_1, d_2) = \frac{ \langle \phi_i (d_1), S_{i,j} \phi_j (d_2) \rangle }{\|\phi_i(d_1)\| \|\phi_j(d_2)\|},$$
where $d_1$ and $d_2$ are documents written in the $i$-th and $j$-th language respectively. If the maximal singular value of $S_{i,j}$ is bounded by $1$, then the similarity scores will lie on the interval $[-1, 1]$. We will provide an overview of the models in Section \ref{sec:models} and then introduce additional notation in \ref{sec:notation}. Starting with Section \ref{sec:kmeans} and ending with Section \ref{sec:hublang} we will describe some approaches to computing $S_{i,j}$ given training data.

\subsection{Cross-Lingual Models}\label{sec:models}
In this section, we will describe several approaches to the problem of computing multilingual similarities introduced in \ref{sec:tfidf}. We present four approaches:
a simple approach based on $k$-means clustering in \ref{sec:kmeans}, a standard approach based on singular value decomposition in \ref{sec:LSI}, a related
approach called Canonical Correlation Analysis (CCA) in \ref{sec:CCA} and finally a new method, which is an extension of CCA to more than two languages in \ref{sec:hublang}.
%
CCA can be used to find correlated patterns for a pair of languages, whereas the extended method optimizes a
Sum of Squared Correlations (SSCOR) between several language pairs, which was introduced in \cite{Kettenring}. The SSCOR problem is difficult to solve in our setting (hundreds of thousands of features, hundreds of thousands of examples). To tackle this, we propose a method which consists of two ingredients.
 The first one is based on an observation that certain datasets (such as Wikipedia) are biased towards one language (English), which can be exploited
 to reformulate a difficult optimization problem as an eigenvector problem. The second ingredient is dimensionality reduction using LSI, which
 makes the eigenvector problem computationally and numerically tractable.
